### 2011-11-07 doc/pldi2012/lowlevel.tex
described valid mask
More edits to the low level

The low-level runtime is a machine-independent layer that provides portability
for our Legion implementation.
%The low-level runtime hides idiosyncracies of the underlying machine while
%at the same time not straying 
%too far from the native capabilities of the underlying hardware.  
The low-level runtime abstracts the machine as a base set of objects:
events, deferred locks, processors, memories, and regions.

@@

The potential space concern with events is addressed by a {\em generation scheme}
that allows events to be reclaimed and reused without tracking references.  The
naive implementation of an event cannot reclaim the storage used by an event
until the event has triggered and all waiters on the event have been
notified.  
%Some of these waiters may take quite a while to query the event.
%Additionaly, we must somehow track how many queries to expect to know
%when they've all been performed.  
Our events include a
{\em generation count} that tracks how many times the event has been triggered,
which allows an event to be reused with a new generation as soon as the previous generation has triggered.
Holders of an older (handle,generation) pair are able to query the newer
incarnation of the event and determine from the generation mismatch that their
event must have already triggered.  With this technique, there is no need to
count references to the event at all. 
% An event can only be reused until the
%generation counter reaches its maximum value, the current implementation
%supports $2^{32}-1$ generations, and should that not be sufficient, extending
%it to $2^{64}-1$ is trivial.

\subsubsection{Deferred Locks}
\label{subsec:defferedlocks}
Events provide a deterministic ordering between operations, 
but the more relaxed coherence do not require strict ordering.  The traditional tool for this job is
the mutex (a.k.a. lock), and it is from the common reader/writer lock that
the low-level runtime's {\em deferred lock} is derived.  The key visible 
difference between a mutex and a deferred lock is that the request does not
block the caller's execution.  Many libraries offer nonblocking locks, but
they achieve this by returning without the lock if it is unavailable.  If
a deferred lock is currently unavailable, the request is guaranteed to be
granted at some time in the future (with no further action required of the 
caller).  That time in the future is described with an event.  Like all other
events, it can be given to other threads, or even stored
in a data structure in memory to be extracted by an arbitrary thread at a 
later time.  This dissociation of a held lock from the thread that
requested it is another important difference between the deferred lock and
mutexes.

@@

\subsubsection{Regions}
\label{subsec:regionmeta}
Application-level regions are broken into three inter-related pieces in the
low-level runtime.  The metadata object tracks global information such as
region size and valid mask.  Allocators are used to dynamically allocate and
free entries in a region.  Finally, instances hold versions of the data for 
a logical region.

>>>>

The low-level runtime is a machine-independent layer that provides portability
for our Legion implementation.
%The low-level runtime hides idiosyncracies of the underlying machine while
%at the same time not straying 
%too far from the native capabilities of the underlying hardware.  
The low-level runtime abstracts the machine as a base set of objects:
events, deferred locks, regions, memories, and processors.

@@

The potential space concern with events is addressed by a {\em generation scheme}
that allows events to be reclaimed and reused without tracking references.  The
naive implementation of an event cannot reclaim the storage used by an event
until the event has triggered and all waiters on the event have been
notified.  
%Some of these waiters may take quite a while to query the event.
%Additionaly, we must somehow track how many queries to expect to know
%when they've all been performed.  
Our events include a
{\em generation count} that tracks how many times the event has been triggered,
which allows an event to be reused with a new generation as soon as the previous generation has triggered.
Holders of an older (handle,generation) pair are able to query the newer
incarnation of the event and determine from the generation mismatch that their
event must have already triggered.  With this technique, there is no need to
count references to the event. 
% An event can only be reused until the
%generation counter reaches its maximum value, the current implementation
%supports $2^{32}-1$ generations, and should that not be sufficient, extending
%it to $2^{64}-1$ is trivial.

\subsubsection{Deferred Locks}
\label{subsec:defferedlocks}
Events provide a deterministic ordering between operations, 
but the more relaxed coherence properties do not require strict ordering.  
The traditional tool for this job is
the mutex (a.k.a. lock), and it is from the common reader/writer lock that
the low-level runtime's {\em deferred lock} is derived.  The key visible 
difference between a mutex and a deferred lock is that the request does not
block the caller's execution.  Many libraries offer nonblocking locks, but
they achieve this by returning without the lock if it is unavailable.  If
a deferred lock is currently unavailable, the request is guaranteed to be
granted at some time in the future (with no further action required of the 
caller).  That time in the future is described with an event.  Like all other
events, it can be given to other threads, or even stored
in a data structure in memory to be extracted by an arbitrary thread at a 
later time.  This dissociation of a held lock from the thread that
requested it is another important difference between the deferred lock and
mutexes.

@@

\subsubsection{Regions}
\label{subsec:regionmeta}
Application-level regions are broken into three inter-related pieces in the
low-level runtime.  The metadata object tracks global information such as
region size and valid mask.  Allocators are used to dynamically allocate and
free entries in a region.  Finally, physical instances hold versions of the data for 
a logical region.
